---
title: "Session12 - Inferential stats - FLER2110"
author: Timo Roettger
date: "8/11/2021"
output: html_document
---

```{r, setup, include=FALSE}

# nifty code using the pacman package
# it checks if the packages specified below are installed, if not, they will be installed, if yes, they will be loaded
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rstudioapi, tidyverse, ggbeeswarm, learnB4SS)

# set the current working directory to the one where this file is
current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_working_dir)

```

# Walkthrough

We will be looking at a familiar data set: the similarity data set from Winter and Matlock.

Let's remind us on whats going on in the data.

```{r data}

# Load our data set
sim <- read_csv("winter_matlock_2013_similarity.csv")

head(sim)

```

Participants drew fictional villages on a map solely based on what they have read about the village (relative to another village). 
The researchers measured the `Distance` of the drawn villages.
Village descriptions and names were either similar or different with respect to their phonological form (`Phon`) or their semantic properties (`Sem`).

## A simple linear model
Let's predict `Distance` based on `Phon` and `Sem`.

```{r lm_simple}

# linear model
xlm <- lm(Distance ~ Phon + Sem,
          data = sim)

summary(xlm)

```

We get a bunch of information. 
Now that we have learned about uncertainty and significance testing, we can interpret the whole model output. 
The model estimates that phonologically similar villages are drawn 5.795 units further away than phonologically different villages (the intercept). 
This difference is statistically not significant (p = 0.1667). 
So if the null hypothesis is true, i.e. phonological similarity has no impact on distance, than the obtained t-value (t = 1.386) or more extreme t-values are too likely for us to reject the null. 
Moreover, the model estimates that semantically similar villages are drawn 10.184 units closer than semantically different villages (the intercept). 
This difference is statistically significant (p = 0.0154). 
So if the null hypothesis is true, i.e. semantic similarity has no impact on distance, than the obtained t-value (t = -2.435) or more extreme t-values are unlikely enough for us to reject the null. 

## Extract predictions
Let's extract the models predictions for all combinations of `Phon` and `Sem`.

```{r predict}

# predict
## new df for prediction
newpreds <- tibble(expand.grid(Phon = unique(sim$Phon),
                               Sem = unique(sim$Sem)))

## predict based on model
sim_preds <- predict(xlm, 
                     newpreds,
                     interval = 'confidence')

sim_predictions <- cbind(newpreds, sim_preds)
sim_predictions

```

## Plot predictions
Now we plot the model estimates and their 95% confidence intervals.

```{r plot}

sim_predictions %>%
  ggplot(aes(x = Phon, 
             y = fit,
             color = Sem)) +
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lwr, 
                    ymax = upr),
                position = position_dodge(0.3),
                width = 0.2) +
  theme_minimal()

```

# Exercises
## Excercise 1

We now go back to the nettle data set introduced at early stages during the class.

```{r data2}

# Load our data set
nettle <- read_csv("nettle_1999_climate.csv")
head(nettle)

```

### (a) 
Run a linear regression that predicts `Population` based on standardized `Area`.
I standardized the predictor `Area` for you.

```{r lm_simple2}

nettle <- nettle %>% 
  mutate(Area_c = (mean(Area) - Area) / sd(Area))

# linear model
xlm_pop <- lm(Population ~ Area_c,
          data = nettle)

summary(xlm_pop)

```

### (b) 
Interpret the model estimates and the p-values

```{r interpret}

# The model estimates that with each unit of Area, Population decreases by 0.38 units.
# This difference is statistically significant (p > 0.00001). 
# So if the null hypothesis was true, i.e. Area did not affect Population, than the obtained t-value (t = -5.834) or more extreme t-values are unlikely enough for us to reject the null. 

```

### (c)
Extract model predictions for `Area_c` and plot a scatter plot with the regression line and the respective 95% confidence interval.

```{r predict2}

# predict
## new df for prediction
newpreds_nettle <- tibble(Area_c = seq(-3, 3, by = 0.1))

## predict based on model
nettle_preds <- predict(xlm_pop, 
                        newpreds_nettle,
                        interval = 'confidence')

nettle_predictions <- cbind(newpreds_nettle, nettle_preds)
nettle_predictions

nettle_area_pop <- 
  nettle_predictions %>%
  ggplot(aes(x = Area_c, 
             y = fit)) +
  geom_text(data = nettle, 
            aes(y = Population, 
                x = Area_c,
                label = Country),
            color = "darkgrey",
            cex = 3) +
  geom_ribbon(aes(ymin = lwr, 
                  ymax = upr),
              fill = 'grey', 
              alpha = 0.5) +
  geom_line() +
  theme_minimal()

nettle_area_pop

```
