---
title: "Homework06"
author: "Markus Heiervang"
date: "19/10/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# nifty code using the pacman package
# it checks if the packages specified below are installed, if not, they will be installed, if yes, they will be loaded
if (!require("pacman")) install.packages("pacman")

pacman::p_load(rstudioapi, tidyverse)

# set the current working directory to the one where this file is
current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path)
setwd(current_working_dir)
```



# Homework (in book: 4.10.1) 
In this exercise, you will perform the analysis corresponding to Figure 4.1 on page 70. 

1. Load in the data set `ELP_frequency.csv` into your R session. 
```{r}
df = read.csv("ELP_frequency.csv")
df
```
2. Use `mutate()` to apply the `log10()` function to the frequency column (`Freq`) (logarithms will be explained in Chapter 5).
```{r}
df = mutate(df, Freq = log10(Freq))
df
```
3. Fit a linear model in which response time (`RT`) are modeled as a function of log frequencies. 
```{r}
# The log10 function is already applied in-place on the dataframe
# So we don't need to apply it again
model = lm(df$RT ~ df$Freq, data = df)
summary(model)
```
4. Create a plot for the relationship between these two variables. 
```{r}
ggplot(df, aes(x = Freq, y = RT)) + geom_point(color = "blue")
```

5. Can you add a horizontal line showing the mean response duration using `geom_hline()` and the `yintercept` aesthetic?
```{r}
ggplot(df, aes(x = Freq, y = RT)) + 
  geom_point(color = "blue") + 
  geom_hline(yintercept = mean(df$RT), color = "red")
```


6. Can you add the regression line from your model?
```{r}
coefs = coefficients(model)
intercept = coefs[["(Intercept)"]]
slope = coefs[["df$Freq"]]

ggplot(df, aes(x = Freq, y = RT)) + 
  geom_point(color = "blue") + 
  geom_hline(yintercept = mean(df$RT), color = "red") + 
  geom_abline(intercept = intercept, slope = slope)
```

(7. Upload your Rmarkdown file (`.Rmd`) and a knitted document (`.html` or `.pdf`) to Canvas. (deadline: 19.10.2021, 23.59h CET))

